---
title: "Tidyverse Assignment - Masculinity Data"
author: "Amber Ferger"
date: "11/28/2019"
output:
  pdf_document: default
  html_document: default
---

## The Data: What Do Men Think It Means to be a Man?
For this assignment, I decided to use the dataset that corresponds to the *What Do Men Think It Means To Be A Man?* article on FiveThirtyEight.com. The article can be found here: https://fivethirtyeight.com/features/what-do-men-think-it-means-to-be-a-man/

It contains the results of a survey of 1,615 adult men conducted by SurveyMonkey in partnership with FiveThirtyEight and WNYC Studios from May 10-22, 2018. 

``` {r, warning = FALSE, message = FALSE}

library(tidyverse)
dat <- as_tibble(read.csv('https://raw.githubusercontent.com/amberferger/DATA607_Masculinity/master/raw-responses.csv'))

```


We have quite a bit of questions in this survey, so we will focus on just a few. For the purpose of this vignette, let's see what role demographics play in the answer to the question **How important is it to you that others see you as masculine?** We'll use the **select** command (from the tidyverse dependency *dplyr*) to return only the columns we are interested in looking at (race and orientation). We'll also use the **filter()** command to subset our data to only individuals that provided a response to these question. 


``` {r}

dat <- dat %>% 
  select(race2,orientation, q0002) %>%
  filter(q0002 != 'No answer' & race2 != 'No answer' & orientation != 'No answer')

```


## Data Aggregation

Our final data set has 1 response variable (the answer to the question) and 2 explanatory variables (our demographic data). We'll use the **group_by** function with the **count()** function to summarize our data. We will then transform our values by creating a percent for each of the of the groupings.
``` {r}

raceCount <- dat %>% 
  group_by(race2, q0002) %>%
  count()

raceCount <- raceCount %>% 
  group_by(race2) %>%
  mutate(RACE_PCT = n/sum(n))

raceCount

```

We'll do the same thing for the **orientation** variable. 

``` {r}

orientationCount <- dat %>% 
  group_by(orientation, q0002) %>%
  count()

orientationCount <- orientationCount %>% 
  group_by(orientation) %>%
  mutate(ORIENTATION_PCT = n/sum(n))

orientationCount

```


## Visualization 

Now let's visualize our data! We'll use the **ggplot** library to take a look: 

``` {r}
library(ggplot2)

ggplot(raceCount, aes(fill=race2, y=RACE_PCT, x=q0002)) + 
    geom_bar(position="dodge", stat="identity") +
    ggtitle("Race vs Answer")

```


``` {r}

ggplot(orientationCount, aes(fill=orientation, y=ORIENTATION_PCT, x=q0002)) + 
    geom_bar(position="dodge", stat="identity") +
    ggtitle("Orientation vs Answer")

```

#### Paul Perez Extension

Extending onto Amber's work above to use two columns in the `group_by()` function to show each orientation by race. Additionally, we'll be able to created barplots and pie charts our of the race and orientation data frames.

## Group By Function using Two Columns
As stated above, one of the ways to view this data is count the number of groupings by grouping the race and orientation of the people surveryed. 


```{r}
race_orientation <- dat %>% group_by(race2, orientation) %>% count()
race_orientation

race_orientation <- race_orientation %>% group_by(orientation) %>% mutate(percentage = n/sum(n))
race_orientation
```


## Creating Bar Plots and Pie Charts

In order to create a pie chart, we'll need to make a barplot first. Using `ggplot`, we can 
```{r}
race_bp <- ggplot(raceCount, aes(x="", y=n, fill=race2))+
geom_bar(width = 1, stat = "identity")
race_bp

orientation_bp <- ggplot(orientationCount, aes(x="", y=n, fill=orientation))+
geom_bar(width = 1, stat = "identity")
orientation_bp
```


```{r}
race_pie <- race_bp + coord_polar("y", start=0) + scale_fill_brewer(palette='Dark1')
race_pie 

orientation_pie <- orientation_bp + coord_polar("y", start=0) + scale_fill_brewer(palette='Dark2')
orientation_pie
```

